Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-15 writeup/gz_hubble_data.tex
Linebreak the prose text at 80 chars

Space-based observatories, particularly the \hubble (\hst), have been responsible for the bulk of imaging studies of high-redshift galaxies. Observations of fixed fields with very deep imaging \citep[eg,][]{wil96,gia04,bec06,dav07,sco07,gro11} give the photometric sensitivity necessary to detect $L^\star$ galaxies at $z>1$, while also providing the angular resolution to distinguish internal structure and characterize the morphology. While these measurements are helped by the fact that the angular diameter distance is relatively flat beyond $z>1$ in a flat $\Lambda$CDM cosmology, the relevant angular scales are only of the order $\sim5-10~\mathrm{kpc}/^{\prime\prime}$ \citep{wri06}. \hst{} can thus resolve much of the structure for a Milky Way-sized galaxy (at least for distinguishing a disk from a bulge), but will be limited for more compact structures. Since the size of galaxies evolves as roughly $r\propto(1+z)^{-1}$ \citep{law12a}, the compact sizes of high-redshift galaxies make detailed morphologies a challenge even for \hst{} \citep{che12}. The public availability of more than $10^5$~galaxies in archival imaging across various studies gives a data sample with the potential for high statistical significance.

@@

%\begin{enumerate}
%    \item Motivation for studying morphology of galaxies
%    \item Particular scientific questions governed by galaxies at $z\sim1$
%    \item Theoretical predictions for galaxy morphology evolution
%    \item Past imaging and morphology studies at $z\sim1$
%    \item Summary of GZ/citizen science work on galaxy morphology to date
%    \item Outline of paper
%\end{enumerate}

@@

\begin{itemize}
\item The All-Wavelength Extended Groth Strip International Survey \citep[AEGIS;][]{dav07} covers a strip centered at $\alpha=14^\textrm{h}17^\textrm{m}, \delta=+52^\circ30^\prime$. This area was selected for a deep survey due to a combination of low extinction and low Galactic/zodiacal emission. The ACS images covered 63 separate tiles over a total area of $\sim710$~arcmin$^2$. The two ACS bands for AEGIS had exposure times of 2300~seconds in F606W (\Vband) and 2100~seconds in F814W (\Iband). The final mosaic images are dithered to a resolution of 0.03~\arcsec/pixel. For extended objects, the limiting magnitudes of sources in AEGIS are 26.23~(AB) in \Vband{} and 25.61 (AB) in \Iband.

@@

\item The Galaxy Evolution from Morphologies and SEDs \citep[GEMS;][]{rix04,cal08} survey is also centered on the Chandra Deep Field-South. The GEMS data covers $\sim800$~arcmin$^2$, completely surrounding the area covered by GOODS-S. Images from ACS in GEMS have 1 orbit per pointing for a total of 63~pointings. The exposure times are 2160 and 2286~seconds in \Vband{} and \zband{}, respectively. The image resolution has a pixel scale of 0.03~\arcsec/pixel. The $5\sigma$ limiting magnitudes for source detection are 25.7 AB in \Vband{} and 24.2~AB in \zband. 
\end{itemize}

@@

%\begin{figure*}
%\center
%\includegraphics[width=160mm]{figures/specz_comparison.pdf}
%\caption{Spectroscopic redshifts from the ACS-GC, 3DHST, and MUSYC catalogs. The number in the upper left of each plot is the percentage of redshifts which agree within $\Delta z < 0.05$ between the two catalogs being compared in each panel. Within this range there is over 98\% agreement in redshifts between all three catalogs.}
%\label{fig:speczs}
%\end{figure*}

%\tabletypesize{\scriptsize}
%\begin{deluxetable*}{l|rr|rrr|rr|r|rr}
%\tablecolumns{12}
%\tablewidth{0pc}
%\tablecaption{GZH redshifts by survey\label{tbl:redshifts}} 
%\tabletypesize{\scriptsize}
%\tablehead{
% &   
%\multicolumn{2}{c}{\underline{ACS-GC}} &
%\multicolumn{3}{c}{\underline{3DHST}} &
%\multicolumn{2}{c}{\underline{MUSYC}} &
%\multicolumn{1}{c}{\underline{UltraVISTA}}&
%\multicolumn{2}{c}{\underline{Total}} 
%\\
%\colhead{Survey} & 
%\colhead{spec-z} & 
%\colhead{photo-z} & 
%\colhead{spec-z} & 
%\colhead{grism-z} & 
%\colhead{photo-z} & 
%\colhead{spec-z} & 
%\colhead{photo-z} & 
%\colhead{photo-z} & 
%\colhead{with redshift} &
%\colhead{in survey}
%}
%\small
%\startdata
%%Survey     Gs        Gp    3ds    3dg     3dp       Ms      Mp       UVp       total
%AEGIS    & 3,656  & 2,941  & 12  &  515  &  249    & 0     & 0     &  0      & 7,373  & 8,507\\
%COSMOS   & 7,201  & 77,435 & 35  &  358  &  26     & 0     & 0     &  2,665  & 85,020 & 92,808 \\
%GEMS     & 387    & 628    & 6   &  99   &  40     & 279   & 7,304 & 0       & 8,743  & 9,304\\
%GOODS-N  & 1,947  & 37     & 418 & 1,545 &  1,381  & 0     & 0     & 0       & 5,328  & 12,030 \\
%GOODS-S  & 1,080  & 4      & 327 & 1,348 &  281    & 816   & 1,184 & 0       & 5,040  & 10,284 \\
%SDSS     & 0      & 0      & 0   &  0    &  0      & 0     & 0     & 0       & 37,545 & 51,861  \\
%\hline
%Total    & 14,271 & 81,045 & 798 & 3,865 & 1,977   & 1,095 & 8,488 & 2,665   & 114,204 & 184,794\\
%\enddata
%\end{deluxetable*}

@@

In the GZH samples, the redshift range is large enough that cosmological evolution of the types and morphologies of galaxies is expected for galaxies in the sample. As a result, the previous methods of correcting for redshift dependent bias do not work. In addition, the effects of band shifting will change the images even more across these redshift ranges. %Figure~\ref{fig:exampleFERENGI} illustrates some of the possible effects of losing features in spiral galaxies at high redshift.

@@

\subsection{Challenges of debiasing questions beyond ``smooth or features''}
\label{ssec:higher_order_tasks}
As with the HST images, each \ferengi{} subject has a varying number of users answering the various questions in the hierarchical decision tree. Every user answers the first question, {\it ``Is the galaxy smooth and rounded, with no sign of a disk?''}; as such the vote fractions \fsmooth, \ffeatures, and \fartifact{} are all computed with the minimum statistical error for any question, with between 40 and 140~responses (see Section~\ref{sec:interface}). The number of users to answer any subsequent question, however, is always equal to or less than the number to answer the preceding question. The average number of responses per task for fourth- or fifth-tier questions such as spiral arm structure (Tasks 12--14) is only $4\pm4$ for the \ferengi{} sample; while this is distribution is strongly bimodal (reflecting the true morphologies of selected galaxies), the very low absolute numbers of votes introduce very high variance when attempting to calculate a statistical correction.

@@

%\begin{figure*}
%\center
%\includegraphics[width=\textwidth]{figures/hubble_z_mu_p_distributions.pdf}
%\caption{Distributions of redshift, surface brightness, and \ffeatures{} for correctable (purple), lower-limit (green), and NEI (pink) galaxies in the full GZH sample. The lower-limit galaxies have higher redshifts, slightly lower surface brightness, and lower values of \ffeatures{} on average than the correctable galaxies. The long tail of NEI galaxies in redshift and surface brightness demonstrates the limits of the \ferengi{} sample, for which there is no data at $z>1$ or $\mu>24$.}
%\label{fig:z_mu_p}
%\end{figure*}

%\subsection{Results of \ferengi{} analysis}
%
%{\bf Old text from Taiwan workshop}
%
%We show here the preliminary results of GZH classifications of images of galaxies placed at artificial redshifts. Figure~\ref{fig:ferengi_results_fake} shows the range of change of vote fractions for the change of \ffeatures{} with redshift, for galaxies with different vote fraction levels, three ranges of surface-brightness levels and 7~evolutionary corrections (this last bit is indicated by the color). 
%
%%-----------------------------------------------------------------------------------------------------------------------------------
%\begin{figure}
%\begin{center}
%
%\includegraphics[width=0.45\textwidth]{figures/somewhat_fake_results.pdf}
%
%\caption{Preliminary results of the \ferengi{} redshifting exercise. WARNING NO USER WEIGHTING.... For a range of vote fraction levels with three surface-brightness levels and 7 evolutionary corrections each (the colors indicate evolutionary correction value with green being $e=0$, we show the range of evolution of the vote fractions for featured vs. smooth with redshift.}
%
%\label{fig:ferengi_results_fake}
%
%\end{center}
%\end{figure}
%%-----------------------------------------------------------------------------------------------------------------------------------

>>>>

Space-based observatories, particularly the \hubble{} (\hst), have been
responsible for the bulk of imaging studies of high-redshift galaxies.
Observations of fixed fields with very deep imaging
\citep[eg,][]{wil96,gia04,bec06,dav07,sco07,gro11} give the photometric
sensitivity necessary to detect $L^\star$ galaxies at $z>1$, while also
providing the angular resolution to distinguish internal structure and
characterize the morphology. While these measurements are helped by the fact
that the angular diameter distance is relatively flat beyond $z>1$ in a flat
$\Lambda$CDM cosmology, the relevant angular scales are only of the order
$\sim5-10~\mathrm{kpc}/^{\prime\prime}$ \citep{wri06}. \hst{} can thus resolve
much of the structure for a Milky Way-sized galaxy (at least for distinguishing
a disk from a bulge), but will be limited for more compact structures. Since
the size of galaxies evolves as roughly $r\propto(1+z)^{-1}$ \citep{law12a},
the compact sizes of high-redshift galaxies make detailed morphologies a
challenge even for \hst{} \citep{che12}. The public availability of more than
$10^5$~galaxies in archival imaging across various studies gives a data sample
with the potential for high statistical significance.

@@

@@

\begin{itemize}

\item The All-Wavelength Extended Groth Strip International Survey
    \citep[AEGIS;][]{dav07} covers a strip centered at
    $\alpha=14^\textrm{h}17^\textrm{m}, \delta=+52^\circ30^\prime$. This area was
    selected for a deep survey due to a combination of low extinction and low
    Galactic/zodiacal emission. The ACS images covered 63 separate tiles over a
    total area of $\sim710$~arcmin$^2$. The two ACS bands for AEGIS had exposure
    times of 2300~seconds in F606W (\Vband) and 2100~seconds in F814W (\Iband). The
    final mosaic images are dithered to a resolution of 0.03~\arcsec/pixel. For
    extended objects, the limiting magnitudes of sources in AEGIS are 26.23~(AB) in
    \Vband{} and 25.61 (AB) in \Iband.

@@

\item The Galaxy Evolution from Morphologies and SEDs
    \citep[GEMS;][]{rix04,cal08} survey is also centered on the Chandra Deep
    Field-South. The GEMS data covers $\sim800$~arcmin$^2$, completely surrounding
    the area covered by GOODS-S. Images from ACS in GEMS have 1 orbit per pointing
    for a total of 63~pointings. The exposure times are 2160 and 2286~seconds in
    \Vband{} and \zband{}, respectively. The image resolution has a pixel scale
    of 0.03~\arcsec/pixel. The $5\sigma$ limiting magnitudes for source
    detection are 25.7 AB in \Vband{} and 24.2~AB in \zband.

\end{itemize}

@@

%\begin{figure*}
%\center
%\includegraphics[width=160mm]{figures/specz_comparison.pdf}
%\caption{Spectroscopic redshifts from the ACS-GC, 3DHST, and MUSYC catalogs.
%The number in the upper left of each plot is the percentage of redshifts which
%agree within $\Delta z < 0.05$ between the two catalogs being compared in each
%panel. Within this range there is over 98\% agreement in redshifts between all
%three catalogs.}
%\label{fig:speczs}
%\end{figure*}

%\tabletypesize{\scriptsize}
%\begin{deluxetable*}{l|rr|rrr|rr|r|rr}
%\tablecolumns{12}
%\tablewidth{0pc}
%\tablecaption{GZH redshifts by survey\label{tbl:redshifts}} 
%\tabletypesize{\scriptsize}
%\tablehead{
% &   
%\multicolumn{2}{c}{\underline{ACS-GC}} &
%\multicolumn{3}{c}{\underline{3DHST}} &
%\multicolumn{2}{c}{\underline{MUSYC}} &
%\multicolumn{1}{c}{\underline{UltraVISTA}}&
%\multicolumn{2}{c}{\underline{Total}} 
%\\
%\colhead{Survey} & 
%\colhead{spec-z} & 
%\colhead{photo-z} & 
%\colhead{spec-z} & 
%\colhead{grism-z} & 
%\colhead{photo-z} & 
%\colhead{spec-z} & 
%\colhead{photo-z} & 
%\colhead{photo-z} & 
%\colhead{with redshift} &
%\colhead{in survey}
%}
%\small
%\startdata
%%Survey     Gs        Gp    3ds    3dg     3dp       Ms      Mp       UVp       total
%AEGIS    & 3,656  & 2,941  & 12  &  515  &  249    & 0     & 0     &  0      & 7,373   & 8,507   \\
%COSMOS   & 7,201  & 77,435 & 35  &  358  &  26     & 0     & 0     &  2,665  & 85,020  & 92,808  \\
%GEMS     & 387    & 628    & 6   &  99   &  40     & 279   & 7,304 & 0       & 8,743   & 9,304   \\
%GOODS-N  & 1,947  & 37     & 418 & 1,545 &  1,381  & 0     & 0     & 0       & 5,328   & 12,030  \\
%GOODS-S  & 1,080  & 4      & 327 & 1,348 &  281    & 816   & 1,184 & 0       & 5,040   & 10,284  \\
%SDSS     & 0      & 0      & 0   &  0    &  0      & 0     & 0     & 0       & 37,545  & 51,861  \\
%\hline
%Total    & 14,271 & 81,045 & 798 & 3,865 & 1,977   & 1,095 & 8,488 & 2,665   & 114,204 & 184,794 \\
%\enddata
%\end{deluxetable*}

@@

In the GZH samples, the redshift range is large enough that cosmological
evolution of the types and morphologies of galaxies is expected for galaxies in
the sample. As a result, the previous methods of correcting for redshift
dependent bias do not work. In addition, the effects of band shifting will
change the images even more across these redshift ranges. 
%Figure~\ref{fig:exampleFERENGI} illustrates some of the possible effects of
%losing features in spiral galaxies at high redshift.

@@

\subsection{Challenges of debiasing questions beyond ``smooth or features''}\label{ssec:higher_order_tasks}

As with the HST images, each \ferengi{} subject has a varying number of users
answering the various questions in the hierarchical decision tree. Every user
answers the first question, {\it ``Is the galaxy smooth and rounded, with no
sign of a disk?''}; as such the vote fractions \fsmooth, \ffeatures, and
\fartifact{} are all computed with the minimum statistical error for any
question, with between 40 and 140~responses (see Section~\ref{sec:interface}).
The number of users to answer any subsequent question, however, is always equal
to or less than the number to answer the preceding question. The average number
of responses per task for fourth- or fifth-tier questions such as spiral arm
structure (Tasks 12--14) is only $4\pm4$ for the \ferengi{} sample; while this
is distribution is strongly bimodal (reflecting the true morphologies of
selected galaxies), the very low absolute numbers of votes introduce very high
variance when attempting to calculate a statistical correction.

@@

%\begin{figure*}
%\center
%\includegraphics[width=\textwidth]{figures/hubble_z_mu_p_distributions.pdf}
%\caption{Distributions of redshift, surface brightness, and \ffeatures{} for
%correctable (purple), lower-limit (green), and NEI (pink) galaxies in the full
%GZH sample. The lower-limit galaxies have higher redshifts, slightly lower
%surface brightness, and lower values of \ffeatures{} on average than the
%correctable galaxies. The long tail of NEI galaxies in redshift and surface
%brightness demonstrates the limits of the \ferengi{} sample, for which there
%is no data at $z>1$ or $\mu>24$.}
%\label{fig:z_mu_p}
%\end{figure*}

%\subsection{Results of \ferengi{} analysis}
%
%{\bf Old text from Taiwan workshop}
%
%We show here the preliminary results of GZH classifications of images of
%galaxies placed at artificial redshifts. Figure~\ref{fig:ferengi_results_fake}
%shows the range of change of vote fractions for the change of \ffeatures{}
%with redshift, for galaxies with different vote fraction levels, three ranges
%of surface-brightness levels and 7~evolutionary corrections (this last bit is
%indicated by the color). 
%
%%---------------------------------------------------------------------------
%\begin{figure}
%\begin{center}
%
%\includegraphics[width=0.45\textwidth]{figures/somewhat_fake_results.pdf}
%
%\caption{Preliminary results of the \ferengi{} redshifting exercise. WARNING
%NO USER WEIGHTING.... For a range of vote fraction levels with three
%surface-brightness levels and 7 evolutionary corrections each (the colors
%indicate evolutionary correction value with green being $e=0$, we show the
%range of evolution of the vote fractions for featured vs. smooth with
%redshift.}
%
%\label{fig:ferengi_results_fake}
%
%\end{center}
%\end{figure}
%%--------------------------------------------------------------------------
